---
title: "Exam 3"
author: "Leah Ellis"
date: "2025-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hi! 

Welcome to my exam 3 submission :D

## Libraries

These are the libraries I used
```{r 0, warning=FALSE,message=FALSE}
library(tidyverse)
library(modelr)
library(easystats)
library(GGally)
library(patchwork)
library(ggpubr)
library(broom)
library(AICcmodavg)
```

## Load and clean the data

Start by reading the data and taking a look at how it is organized.
```{r 1}
df_raw <- read.csv("FacultySalaries_1995.csv")
glimpse(df_raw)
```
Alright... this looks like a mess but I cannot quite tell how to clean this just yet...
All data types are correct, so it seems.
The FacultyAll seems to just be a sum of all the other num columns, so that could be considered a bit extra- but I don't know if removing it would actually help it be more readible. There's a chance that there is faculty included in the total sum that are not profs or instructors- it's a little vague in that way.

Wait... maybe I see one way to fix this mess.

Let's start by removing these pesky combined variables.
```{r 2}
df <- df_raw %>% select(-AvgProfSalaryAll, -AvgProfCompAll, -NumFacultyAll)

```

Then let's make some hefty use of pivot_longer and chaining some joining together.
```{r 3}

titlePattern='(Instructor|FullProf|AssocProf|AssistProf)'

df1 <- df %>% select(FedID, UnivName, State, Tier, NumFullProfs, NumAssocProfs, NumAssistProfs, NumInstructors) %>% 
  pivot_longer(cols = starts_with("Num"), names_to = "Title", names_pattern = titlePattern, values_to = "Count")

df2 <- df %>% select(FedID, UnivName, State, Tier, AvgFullProfComp, AvgAssocProfComp, AvgAssistProfComp) %>% 
  pivot_longer(cols=contains("Comp"), names_to = "Title", names_pattern = titlePattern, values_to = "Avg_Comp")

df3 <- df %>% select(FedID, UnivName, State, Tier, AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary) %>% 
  pivot_longer(cols=contains("Salary"), names_to = "Title", names_pattern = titlePattern, values_to = "Avg_Salary")

by <- join_by(FedID, UnivName, State, Tier, Title) 
df4 <- full_join(df1, df2, by)
df4 <- full_join(df4, df3, by)
```


## Recreate the graph

Now that the data is manageable, it's time to re-create the graph.

```{r 4, warning=FALSE}

df_g <- df4 %>%
  filter(Tier == "I" | Tier == "IIA" | Tier == "IIB") %>% 
  mutate(Title = case_when(
    Title == "AssistProf" ~ "Assist",
    Title == "AssocProf" ~ "Assoc",
    Title == "FullProf" ~ "Full"
  )) %>%
  filter(!is.na(Title)) %>% 
  rename(Salary = Avg_Salary) %>% 
  rename(Rank = Title)

df_g %>% 
  ggplot(aes(x=Rank, y=Salary, fill=Rank)) +
  facet_wrap(vars(Tier)) +
  geom_boxplot(alpha=1) +
  theme_minimal() 
```
That's pretty much exact, I'd say.

## ANOVA

```{r anova1}
aov1 <- df4 %>% aov(data=., Avg_Salary ~ Tier + Title)
summary(aov1)
```
Running it through a two-way ANOVA, we can see that there is at least a small correlation from both Tier and Faculty Rank/Title to the avg_salary of any member of faculty.

## Juniper Oils

First we take a look at, and then clean the data.
Raw Data:
```{r j1}
jo_df_raw <- read.csv("Juniper_Oils.csv")
glimpse(jo_df_raw)
```

Clean:
```{r j2, message=FALSE, warning=FALSE, error=FALSE}
chemlist <- c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene","alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene","cuparene","compound 1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound 2","thujopsenal")

chemlist2 <- c("alpha.pinene","para.cymene","alpha.terpineol","cedr.9.ene","alpha.cedrene","beta.cedrene","cis.thujopsene","alpha.himachalene","beta.chamigrene","cuparene","compound.1","alpha.chamigrene","widdrol","cedrol","beta.acorenol","alpha.acorenol","gamma.eudesmol","beta.eudesmol","alpha.eudesmol","cedr.8.en.13.ol","cedr.8.en.15.ol","compound.2","thujopsenal")

jo_df <- jo_df_raw %>% 
  select(SampleID, YearsSinceBurn, chemlist2) %>% 
  pivot_longer(cols = chemlist2, names_to = "ChemicalID", values_to = "Concentration")
  
glimpse(jo_df)

```

Then, I replicate the plot.
```{r j3, warning=FALSE}

jo_df %>% 
  ggplot(aes(x=YearsSinceBurn, y=Concentration)) + 
  facet_wrap(vars(ChemicalID), scales = "free") +
  geom_smooth() +
  theme_minimal()

```
## GLM Models

Now I'll make a generalized linear model to find which chemicals that are P < 0.05 affected by "Years Since Burn" 

```{r j4}

mod1 <- jo_df %>% glm(data=., Concentration ~ YearsSinceBurn + ChemicalID)

tidy(mod1) %>%
  mutate(term = str_remove(term, '^ChemicalID')) %>% 
  filter(p.value < 0.05) %>% 
  arrange(term) %>% 
  print()

```

## Conclusion

All done, hope it's okay it's about twenty minutes late at this point :p
I submitted a less complete version of this right before the clock rung twelve, this finished version I imagine could be treated as a resubmission... which I imagine aren't allowed. Ah well, it was fun to finish it at least. I like broom I think.

Cya!

  - Leah Ellis
